---
title: "N. Umpqua 2024 Daily Temp Stats & Air/Water Regressions"
author: "J.Hart"
date: "2025-01-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


For each stream temperature site (including USGS sites and our temperature logger sites), please summarize:
1) maximum daily maximum stream temperature (ie, just the warmest temperature recorded at each site).
2) the number of days where daily maximums exceeded 20 C.
3) linear regressions of daily maximum stream temperatures on daily maximum air temperatures for the June 1 -  Aug 30 period. Use the air temperature data from the NorthUmpAtSteamboat site  (logger # 21679162). We will want to take the slope coefficients and R squared values from each regression. 

I think we should summarize all this in a table in the manuscript, so providing the info in excel table form is the easiest for me to deal with. I also think it would be good to output plots for all of the air stream temperature regressions to visually check for fit and outliers, etc.. Right now I would envision this as supplementary material, but not necessarily in the main manuscript.


```{r}
# Load packages

library(tidyverse)
library(readxl)
library(janitor)
library(writexl)

```


```{r}
# Read in logger metadata

log_metadata <- read_xlsx("NUmpTemperature2024.xlsx") %>% 
 clean_names()

# Create data frame for logger lat/long

logger_loc <- log_metadata %>% 
  select(logger_id, lat, long)

# Create date frame for logger site location descriptions

logger_site <- log_metadata %>% 
  select(logger_id, site)

# Create empty data frame for combined air/water daily stats

mat = matrix(ncol = 0, nrow = 0) 
daily_air_master = data.frame(mat)
daily_water_master = data.frame(mat)

# Read in each logger table & merge metadata info

temp <- read_xlsx("2024 Logger Sites/Water/21679182.xlsx") %>% 
 clean_names() %>% 
 mutate(logger_id = "21679182",
        type = "Water") %>% 
 select(-number) %>% 
 rename("date"= "date_time_pdt",
        "temp_c"= "temperature_c") %>% 
 separate('date',
          into = c('longdate', 'time'),
          sep= ' ') %>% 
 separate('longdate',
                  into = c('year', 'month', 'day'),
                  sep= '-',
                  remove = FALSE) 

temp <- merge(temp, logger_loc, by="logger_id") 

# Calcuate daily temp statistics 

daily_temp_stats <- temp %>%
  group_by(logger_id, type, year, month, day, longdate, lat, long) %>%
  summarise(mean_temp = round(mean(temp_c), digits=2),
            min_temp = round(min(temp_c), digits=2),
            max_temp = round(max(temp_c), digits=2))
  
# Merge individual tables to master table after each run
  
daily_air_master <- rbind(daily_air_master, daily_temp_stats)
daily_water_master <- rbind(daily_water_master, daily_temp_stats)

```


```{r USGS Station Stats}

# Process USGS Gauge site tables individually & merge to master

temp <- read_xlsx("USGS Sites/USGS Site 14317450 (Idleyld Park).xlsx") %>% 
 clean_names() %>% 
 mutate(type = "Water") %>% 
rename("longdate" = "datetime",
       "mean_temp" = "mean_c",
       "min_temp" = "min_c",
       "max_temp" = "max_c") %>% 
separate('longdate',
                  into = c('year', 'month', 'day'),
                  sep= '-',
                  remove = FALSE) 

temp <- merge(temp, logger_site, by="logger_id")

daily_water_master <- rbind(temp, daily_water_master)
```


```{r}
# Merge location descriptions to master

daily_air_master <- merge(daily_air_master, logger_site, by="logger_id") 
daily_water_master <- merge(daily_water_master, logger_site, by="logger_id") 

# Write daily temp stats master to table

write_xlsx(daily_air_master, "2024 Logger Sites/Air/daily_air_master.xlsx")
write_xlsx(daily_water_master, "2024 Logger Sites/Water/daily_water_master.xlsx")
```


```{r}
# Review prelim daily air temp plots
a <- ggplot(data = daily_air_master, aes(x = as.Date(longdate), y = mean_temp)) + geom_line()
a + facet_wrap(~logger_id)

# Review prelim daily water temp plots
w <- ggplot(data = daily_water_master, aes(x = as.Date(longdate), y = mean_temp)) + geom_line()
w + facet_wrap(~logger_id)
```



```{r MDMS & Days Above 20C}

# Maximum daily maximum temperature

MDMT <- daily_water_master %>% 
  filter(longdate > "2024-06-19" & longdate <= "2024-09-15") %>% 
  group_by(logger_id, site) %>% 
  summarize(mdms = max(max_temp))

# Days maximum temperature above 20C

DaysAbove20 <- daily_water_master %>% 
  filter(longdate > "2024-06-19" & longdate <= "2024-09-15") %>% 
  filter(max_temp > 20) %>% 
  group_by(logger_id,site) %>% 
  summarize(daysabove20 = n())
  
```



## Max temp air/water correlations

```{r}
# Prepare air logger data for regression

air <- daily_air_master %>% 
  filter(logger_id == "21679162") %>% 
  mutate(longdate = make_date(year,month,day)) %>% 
  filter(longdate > '2024-06-19' & longdate <= '2024-09-15') %>% 
  mutate(air_max = max_temp)

# Prepare individual water logger data for regression

water <- daily_water_master %>% 
  filter(logger_id == '21679164') %>% 
  mutate(longdate = make_date(year,month,day)) %>%
  filter(longdate > '2024-06-19' & longdate <= '2024-09-15') %>% 
  mutate(water_max = max_temp) 

# Join table for linear regression
airwater_join <- water %>% 
  left_join(air, by = "longdate")
```


```{r}
# Create equation formula for regression equation

lm_eqn <- function(df){
    m <- lm(water_max ~ air_max, df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 3),
              b = format(unname(coef(m)[2]), digits = 3),
             r2 = format(summary(m)$adj.r.squared, digits = 3)))
    as.character(as.expression(eq));
}                 


# Calculate Pearson correlation coefficient
correlation <- cor(airwater_join$water_max, airwater_join$air_max, method = 'pearson')
correlation

fit <- lm(water_max ~ air_max, data=airwater_join)
summary(fit)
```


```{r}
# Plot linear regression

a <- ggplot(airwater_join, aes(x=air_max, y = water_max)) +
  geom_point(color='blue') +
  geom_smooth(method = "lm", se = TRUE, colour = 'black') +
  theme_light() + 
 labs(title="Steamboat Cr. Above Canton",
       subtitle = "6/20/2024 – 9/15/2024",
       x="Daily Max Air temperature (°C)",
       y="Daily Max Water temperature (°C)") #+ 
#geom_text(x = 20, y = 16, label = lm_eqn(airwater_join), parse = TRUE)  + 
#geom_text(x = 20, y = 15, label = paste("Pearson R:", round(correlation,3)))

a
```

```{r}
# Export regression plot
ggsave("2024 Logger Sites/Water/Regression Plots/Max Daily/Steamboat Cr. Above Canton.png", width = 10, height = 7)
```


## Mean temp air/water correlations

```{r}
# Prepare air logger data for regression

air <- daily_air_master %>% 
  filter(logger_id == "21679162") %>% 
  mutate(longdate = make_date(year,month,day)) %>% 
  filter(longdate > '2024-06-19' & longdate <= '2024-09-15') %>% 
  mutate(air_mean = mean_temp)

# Prepare individual water logger data for regression

water <- daily_water_master %>% 
  filter(logger_id == '21679164') %>% 
  mutate(longdate = make_date(year,month,day)) %>%
  filter(longdate > '2024-06-19' & longdate <= '2024-09-15') %>% 
  mutate(water_mean = mean_temp) 

# Join table for linear regression
airwater_join <- water %>% 
  left_join(air, by = "longdate")
```

```{r}
lm_eqn <- function(df){
    m <- lm(water_mean ~ air_mean, df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 3),
              b = format(unname(coef(m)[2]), digits = 3),
             r2 = format(summary(m)$adj.r.squared, digits = 3)))
    as.character(as.expression(eq));
}                 


# Calculate Pearson correlation coefficient
correlation <- cor(airwater_join$water_mean, airwater_join$air_mean, method = 'pearson')
correlation

fit <- lm(water_mean ~ air_mean, data=airwater_join)
summary(fit)

```

```{r}
# Plot linear regression

b <- ggplot(airwater_join, aes(x=air_mean, y = water_mean)) +
  geom_point(color='blue') +
  geom_smooth(method = "lm", se = TRUE, colour = 'black') +
  theme_light() + 
 labs(title="Steamboat Cr. Above Canton",
       subtitle = "6/20/2024 – 9/15/2024",
       x="Daily Mean Air temperature (°C)",
       y="Daily Mean Water temperature (°C)") #+
#  geom_text(x = 20, y = 16, label = lm_eqn(airwater_join), parse = TRUE) 
#  geom_text(x = 20, y = 15, label = paste("Pearson R:", round(correlation,3)))

b
```

```{r}
# Export regression plot
ggsave("2024 Logger Sites/Water/Regression Plots/Mean Daily/Steamboat Cr. Above Canton.png", width = 10, height = 7)
```

